### 2015-12-10 paper/4-performance.tex
Added a link to the repository; slightly more relevant info on the used CPU
Wrote explanation of figures for sequential simulation; thinking about having different plots

For all benchmarks, results were all well within 1\% deviation of the average, such that only the average is used in the remainder of this section.
The same compilation flags were used for both adevs and dxex benchmarks (``\texttt{-O3}'').
To guarantee comparable results, no IO was performed during the benchmarking phase.
Simulation traces were used to verify that both adevs and dxex models have exactly the same behaviour.

@@

\begin{figure}
    \center
    %\includegraphics[width=\columnwidth]{fig/queue_model.eps}
    \caption{Queue model for depth 3 and width 2.}
    \label{fig:queue_model}

    %\includegraphics[width=\columnwidth]{fig/interconnect_model.eps}
    \caption{HighInterconnect model for four models.}
    \label{fig:interconnect_model}

    %\includegraphics[width=\columnwidth]{fig/PHOLD_model.eps}
    \caption{PHOLD model for four models, split over two nodes. In parallel simulation, each coupled model is simulated at a different node.}
    \label{fig:PHOLD_model}
\end{figure}

\subsection{Sequential Simulation}
Despite our core contribution being mainly in the parallel simulation, we still value a comprehensive comparison of sequential simulation results.
First, and foremost, as parallel simulation results are tightly linked to the sequential simulation results: parallel simulation merely adds a synchronization layer over different, essentially sequential, simulation kernels.
Second, since parallel simulation results are validated through the use of adevs.
To provide a more comprehensive comparison to adevs in the parallel simulation benchmarks, sequential simulation results need to be compared.

Benchmarks have been run on an \textit{Intel i7-2670QM} with \textit{16GB of main memory} on Linux.
Results were obtained through the use of \textit{perf}.
A selection of call-graphs and more detailed results can be found online.

%Note that a fixed timeadvance (e.g. devstone) is set at 100.
\subsubsection{Devstone}
The Devstone \cite{DEVStone} benchmark is highly hierarchical. As can be seen in Figure~\ref{fig:Devstone}, dxex is outperforming adevs considerably. This can be explained by the fact that, using Direct connection, the hierharchical structure is being flattened whereas, in the adevs case, considerable overhead is caused by traversing the coupled model's structure to pass events.
\begin{figure}[h]
	\includegraphics[width=.5\textwidth]{fig/fig1.eps}
	\caption{Devstone}
	\label{fig:Devstone}
\end{figure}

\subsubsection{PHold}
PHold \cite{PHOLD} is a benchmark that focuses in particular on parallel execution; the sequential runtime is therefore only added as a baseline. A model selects the destination of a message at runtime using a state-saved RNG, which takes up a non-trivial amount of runtime. The results of the benchmark are visualized in Figure~\ref{fig:PHold}.
%% needs further clarification...
\begin{figure}[h]
	\includegraphics[width=.5\textwidth]{fig/fig4.eps}
	\caption{PHold}
	\label{fig:PHold}
\end{figure}

\subsubsection{Interconnect}
Interconnect~\cite{MasterThesis} is a benchmark where all models broadcast, creating a complete graph in dependencies between models. As the model count increases, we see (see Figure~\ref{fig:Interconnect} the expected quadratic increase in runtime for both adevs and dxex, but an increasing penalty for dxex. Profiling shows this is entirely due to the heap allocation of messages, which even though minimized by using memory pools remains significant.
\begin{figure}[h]
	\includegraphics[width=.5\textwidth]{fig/fig3.eps}
	\label{fig3.eps}
	\caption{Interconnect}
	\label{fig:Interconnect}
\end{figure}

\subsection{Parallel Simulation}
%TODO explain that we will now compare different synchronization protocols
%TODO compare different synchronization protocols with adevs' conservative, showing that we are faster (or at least, not that much slower) than adevs
By default, the benchmarks use 4 kernels for parallel simulation.
\subsubsection{Devstone}
The flattened models are allocated to kernels by giving each kernel a distinct section of the chain, resulting in a low ratio of inter-kernel to intra-kernel messages. For the optimistic case, this can cause more reverts since the kernels will start to drift faster as the model-count increases. Furthermore, optimistic synchronization is quite sensitive to an increase in kernels, since the delay before a revert propagates, increases. This benchmark requires a specific warm-up time; for $n=d\times w$ models, it takes $\textup{timeadvance}()*(n-1)$ transitions to activate the last model in the chain. When using parallel processing, this can be reduced to $\textup{timeadvance}()*(n\frac{kernels-1}{kernels} -1)$ before the last kernel becomes active.\\
Figure~\ref{fig:DevstoneParallel} shows that dxex outperformance adevs (in both optimistic and conservative case).
\begin{figure}
	\includegraphics[width=.5\textwidth]{fig/fig2.eps}
	\label{fig:DevstoneParallel}
	\caption{DevStone parallel}
\end{figure}
\subsubsection{PHold}
%Phold
In Phold, the allocation is specified in the benchmark itself. Each kernel manages a single node with a constant set of sub-nodes. The parameter $R$ (which is here set to $10$) determines the percentage of remote destination models.\\
The dynamic dependency graph is a very sparse version of the static dependency graph, penalizing the conservative case. The lookahead is $\epsilon$, so the conservative case spends most of its time crawling in steps of $\epsilon$. Since the dependency graph between kernels is a complete graph, this is not a simulation that scales in our implementation. For $N$ kernels, each kernel has to query the null-time of $N-1$ kernels, resulting in O($N^2$) polling behaviour. In a non-cyclic simulation with a non-trivial lookahead (like in, e.g., Devstone), that choice does pay off (see Figure~\ref{fig:DevstoneParallel}). Adevs's lesser performance is due in part to their lookahead management, which after profiling shows to spend a non-trivial amount of time in exception handling code.\\
The optimistic case suffers little from the above problems; due to the high interconnectivity, however, a cascading revert is still possible. With the percentage of remotes equal to $100$, Phold reflects interconnect in behaviour, which is why the $R$ value is not dimensioned here. A revert is very expensive in PHold due to our usage of C++11's random nr generators. The cost of a revert is dominated by the recalculation of destination models, not in allocating/deallocating states/messages. Once a revert happens the drift between kernels increases fast, increasing the likelihood of more reverts. Despite all this, optimistic can for low $R$ values quickly exploit the uncertainty that slows down conservative in this benchmark.

>>>>

For all benchmarks, results were all well within 1\% deviation of the average, such that only the average is used in the remainder of this section.
The same compilation flags were used for both adevs and dxex benchmarks (``\texttt{-O3}'').
To guarantee comparable results, no IO was performed during the benchmarking phase.
Simulation traces were used to verify that both adevs and dxex models have exactly the same behaviour.
All benchmarks were performed using Linux, but our simulation tool works equally well on Windows and Mac.

@@

\begin{figure}
    \center
    %TODO
    %\includegraphics[width=\columnwidth]{fig/queue_model.eps}
    \caption{Queue model for depth 3 and width 2.}
    \label{fig:queue_model}

    %TODO
    %\includegraphics[width=\columnwidth]{fig/interconnect_model.eps}
    \caption{HighInterconnect model for four models.}
    \label{fig:interconnect_model}

    %TODO
    %\includegraphics[width=\columnwidth]{fig/PHOLD_model.eps}
    \caption{PHOLD model for four models, split over two nodes. In parallel simulation, each coupled model is simulated at a different node.}
    \label{fig:PHOLD_model}
\end{figure}

\subsection{Sequential Simulation Execution Time}
Despite our core contribution being mainly in the parallel simulation, we still value a comprehensive comparison of sequential simulation results.
First, and foremost, as parallel simulation results are tightly linked to the sequential simulation results: parallel simulation merely adds a synchronization layer over different, essentially sequential, simulation kernels.
Second, since parallel simulation results are validated through the use of adevs.
To provide a more comprehensive comparison to adevs in the parallel simulation benchmarks, sequential simulation results need to be compared.
Only the Queue and HighInterconnect models are relevant for sequential simulation.

\subsubsection{Queue}
In the Queue model, we increase both the width and depth simultaneously, causing a quadratic growth in the number of atomic models.
As can be seen in Figure~\ref{fig:Queue_benchmark}, dxex considerably outperforms adevs.
Through careful analysis of profiling results, we determined that adevs spends much time in handling simulation messages, whereas this is mostly avoided due to the differently designed simulation algorithm of dxex.
Both simulation tools have quadratically increasing execution times, though dxex is much faster thanks to its more efficient simulation control algorithms.

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{fig/fig1.eps}
	\caption{Queue benchmark results}
	\label{fig:Queue_benchmark}
\end{figure}

\subsubsection{HighInterconnect}
In the HighInterconnect model, we increase the number of atomic models, thus quadratically increasing the number of couplings.
As can be seen in Figure~\ref{fig:HighInterconnect_benchmark}, adevs now outperforms dxex by a fair margin.
Analysis showed that this slowdown is caused by the high amount of exchanged events.
Event creation is found to be much slower in dxex than it is in adevs, even despite the use of memory pools in dxex.

%TODO only include sequential results?
\begin{figure}[h]
	\includegraphics[width=.5\textwidth]{fig/fig3.eps}
	\label{fig3.eps}
	\caption{Interconnect}
	\label{fig:Interconnect}
\end{figure}

\subsection{Parallel Simulation Execution Time}
%TODO explain that we will now compare different synchronization protocols
%TODO compare different synchronization protocols with adevs' conservative, showing that we are faster (or at least, not that much slower) than adevs
By default, the benchmarks use 4 kernels for parallel simulation.
\subsubsection{Devstone}
The flattened models are allocated to kernels by giving each kernel a distinct section of the chain, resulting in a low ratio of inter-kernel to intra-kernel messages. For the optimistic case, this can cause more reverts since the kernels will start to drift faster as the model-count increases. Furthermore, optimistic synchronization is quite sensitive to an increase in kernels, since the delay before a revert propagates, increases. This benchmark requires a specific warm-up time; for $n=d\times w$ models, it takes $\textup{timeadvance}()*(n-1)$ transitions to activate the last model in the chain. When using parallel processing, this can be reduced to $\textup{timeadvance}()*(n\frac{kernels-1}{kernels} -1)$ before the last kernel becomes active.\\
Figure~\ref{fig:DevstoneParallel} shows that dxex outperformance adevs (in both optimistic and conservative case).
\begin{figure}
	\includegraphics[width=.5\textwidth]{fig/fig2.eps}
	\label{fig:DevstoneParallel}
	\caption{DevStone parallel}
\end{figure}
\subsubsection{PHold}
%Phold
In Phold, the allocation is specified in the benchmark itself. Each kernel manages a single node with a constant set of sub-nodes. The parameter $R$ (which is here set to $10$) determines the percentage of remote destination models.\\
The dynamic dependency graph is a very sparse version of the static dependency graph, penalizing the conservative case. The lookahead is $\epsilon$, so the conservative case spends most of its time crawling in steps of $\epsilon$. Since the dependency graph between kernels is a complete graph, this is not a simulation that scales in our implementation. For $N$ kernels, each kernel has to query the null-time of $N-1$ kernels, resulting in O($N^2$) polling behaviour. In a non-cyclic simulation with a non-trivial lookahead (like in, e.g., Devstone), that choice does pay off (see Figure~\ref{fig:DevstoneParallel}). Adevs's lesser performance is due in part to their lookahead management, which after profiling shows to spend a non-trivial amount of time in exception handling code.\\
The optimistic case suffers little from the above problems; due to the high interconnectivity, however, a cascading revert is still possible. With the percentage of remotes equal to $100$, Phold reflects interconnect in behaviour, which is why the $R$ value is not dimensioned here. A revert is very expensive in PHold due to our usage of C++11's random nr generators. The cost of a revert is dominated by the recalculation of destination models, not in allocating/deallocating states/messages. Once a revert happens the drift between kernels increases fast, increasing the likelihood of more reverts. Despite all this, optimistic can for low $R$ values quickly exploit the uncertainty that slows down conservative in this benchmark.
